hbm_paper_results

### 2024-05-24 paper_main_results.Rmd
cleaning up, re-adding plot of sex-effects in variance

sex.var.plot <- ggplot(sex.var.plot.df) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = median_sex_diff)) +
  #scale_fill_gradient2() +
  # ggseg::geom_brain(data=med_wall, atlas = dk, 
  #            position = position_brain(hemi ~ side),
  #            aes(fill=median_sex_diff)) +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.25, 0.25)) #
saveRDS(sex.var.plot, file="figs/sex_var_plot.RDS")
```

@@

sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_sex_diff),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.25, 0.25), name = "Male - Female Median \nCentile Error") #,
# sex.var.plot.global

#save RDS
saveRDS(sex.var.plot.global, file="figs/sex_var_plot_global.RDS")
```

Combine plots - need to keep playing around to make the spacing pretty when it saves out
```{r}
p_sex <- ggarrange(
  sex.var.plot.global, sex.var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

@@

>>>>

sex.var.plot <- ggplot(sex.var.plot.df) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = median_sex_diff)) +
  #scale_fill_gradient2() +
  # ggseg::geom_brain(data=med_wall, atlas = dk, 
  #            position = position_brain(hemi ~ side),
  #            aes(fill=median_sex_diff)) +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.25, 0.25)) #
```

@@

sex.var.plot.global <- ggplot(sex.var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_sex_diff),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.25, 0.25), name = "Male - Female Median \nCentile Error") #,
# sex.var.plot.global
```

Combine plots - need to keep playing around to make the spacing pretty when it saves out
```{r}
p_sex <- ggarrange(
  sex.var.plot.global, sex.var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.2),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

@@

# SCALE EFFECTS OF SEX

Get differences in male vs female predicted variances at mean sample age

```{r}
var.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

for (file in var.csvs) {
  # Read each CSV file
  data <- fread(file)

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = gsub("_data|_variance.csv", "", Source_File),
           sex_standard = sex_effect/f.var)

  # Bind the data to the combined dataframe
  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
}
range(combined_var$sex_standard)

combined_var <- combined_var %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

#get summary df for p vals, plotting
drop1.df <- fread(paste0(data_path, "/ukb_basic/drop1_tests.csv"), stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  rename(dataset=Dataset,
         pheno=Model) %>%
    dplyr::filter(Moment == "sigma" & Term == "sexMale") %>%
  mutate(dataset = gsub("-data|ukb-CN-data-simsites-", "", dataset)) %>% #clean up dataset naming
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "raw"))

sigma.sex.df <- left_join(combined_var, drop1.df)
sigma.sex.df2 <- left_join(sigma.sex.df, dk.parc)

#combined_var <- Reduce(function(x, y) full_join(x, y), list(combined_var, dk.parc, drop1.df))
```

```{r}
dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
#var plot
var.plot.df <- sigma.sex.df2 %>%
  dplyr::filter(pheno_cat != "Global Volume" & dataset != "cf") %>%
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         sex_standard = ifelse(drop1.pval <0.05, sex_standard, NA)) %>%
  group_by(dataset, pheno_cat)

range(var.plot.df$sex_standard, na.rm=T)

tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="raw")
tp$sex_standard <- 1

var.plot <- ggplot(var.plot.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_standard)) +
  #scale_fill_gradient2() +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
var.plot

#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
saveRDS(var.plot, file="figs/sex_var_plot.RDS")
```

```{r}
#plot global
var.plot.global.df <- sigma.sex.df2 %>%
  dplyr::filter(pheno_cat == "Global Volume" & dataset != "cf") %>%
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  dplyr::mutate(tissue_class = case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA_character_),
         sex_standard = ifelse(drop1.pval <0.05, sex_standard, NA)) %>%
  group_by(dataset)

tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="raw")
tp$sex_standard <- 1

range(var.plot.global.df$sex_standard, na.rm=T)

var.plot.global <- ggplot(var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = sex_standard),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
#var.plot.global
saveRDS(var.plot.global, file="figs/sex_var_plot_global.RDS")
```

Combine plots
```{r}
p <- ggarrange(
  var.plot.global, var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )
```
